---
title: "Lab1"
author: "Anna Matarazzo"
date: "2/21/2021"
bibliography: export-data-2.bib
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Written QALMRI

This is a QALMRI for @slamecka_retroactive_1960.


### Q

_Broad question_: Is Retroactive Inhibition (RI) sensitive to connected speech?

_Specific question_: Will manipulating the degree of practice have an effect on the recall of a text in the same way it has on the unconnected random list of words?

### A

_Hypothesis_: 

+	recall of a prose text is a subject to RI.
+	the number of recalled words is a direct function of amount of rehearsal in Original Learning (OL).
+	the number of recalled words is an inverse function of rehearsal in Interpolated Learning (IL).

_Alternative hypothesis_: 

+	recall of a prose text is not a subject to RI.
+	the number of recalled words is not a direct function of amount of rehearsal in OL. 
+ the number of recalled words is not an inverse function of rehearsal in IL.

### L

-	if the mean number of correct anticipations increases with the number of trials of OL, then the number of recalled words is a direct function of amount of rehearsal in OL; 

- if the mean number of correct anticipations does not increase with the number of trials of OL, then the number of recalled words is not a function of amount of rehearsal in OL.

-	if the number of correct anticipations decreases as the number of IL trials increases, then the number of recalled words is an inverse function of rehearsal in IL;

- if the number of correct anticipations does not decrease as the number of IL trials increases, then the number of recalled words is not an inverse function of rehearsal in IL.


### M

36 people were divided into 9 groups depending on the number of OL (2, 4, 8) and IL (0, 4, 8) trials.

Passages of text were grouped in 4 pairs – 1st and 3rd in OL, 2nd and 4th in IL, counterbalanced. 

Sessions began with color-guessing. Then a number of OL and IL trials happened based on the assigned groups. 

After a 9.5-minute break, participants were given 4 relearning sessions on OL, followed by 3 relearning sessions on IL.

Independent and dependent variables (IV and DV): IV-1 – number of OL trials, IV-2 – number of IL trials, DV – number of correct anticipations. 



### R

Increasing the number of OL at every level of IL produced the increase in the mean number of correct anticipations.

IL means increased with increasing OL trials.

Recall as a function of OL is approximately a linear function.

Regardless of the number of OL, IL-0 group recalled more words than IL-4 and IL-8 groups, thus RI was obtained.

At IL-4 and IL-8, RI percent went down with the increase in OL. RI percent decreased as the number of IL went up.


### I

Recall of a text written in prose was a subject of RI. 

OL and IL acquisition was a function of number of practice trials.

Recall was a direct function of the number of OL trials and an inverse function of the number of IL trials.

One possible confound of the data collection process might have been the fact that participants in groups IL-0 and IL-4 needed to fill in the temporal gap by getting involved into “rest activity” of color-guessing. 

There is a possibility that if the same was true for OL-2 and OL-4 groups (this is not clear from the paper), then the decreased number of correct anticipators as a function of number of OL trials could have been a function of increased time or increased number of “color guessing activities”.




## Coding Problem

1. Create a dataframe that capable of representing the design, in as complete a way as you can, in the @slamecka_retroactive_1960 paper.

```{r}

n <- 36
OL <- 3
IL <- 3

design <- matrix(4,ncol=3, nrow=3)
colnames(design) <- c("OL_2","OL_4","OL_8")
row.names(design) <- c("IL_0","IL_4","IL_8")

subjects <- 1:36

OL <- rep(c(2,4,8),12)
IL <- rep(c(0,4,8), each=12)
recall <- sample(0:20, 36, replace=TRUE)

library(tibble)

long_design <- tibble(subjects,
                       OL,
                       IL,
                      recall)

library(dplyr)
long_design %>%
  group_by(OL,IL) %>%
  count()

# mixed design

mixed_design <- tribble(
~Subjects, ~OL, ~IL,
#--|--|----
1,2,0,
1,4,4,
1,8,8,
2,4,0,
2,8,4,
2,2,8,
3,8,0,
3,2,4,
3,4,8,
4,2,4,
4,4,0,
4,8,8,
5,4,4,
5,2,8,
5,8,0,
6,8,4,
6,4,8,
6,2,0,
7,2,8,
7,4,0,
7,8,4,
8,4,8,
8,2,4,
8,8,0,
9,8,8,
9,4,4,
9,2,0
)

mixed_design %>%
  group_by(OL,IL) %>%
  count()

full_mixed_design <- rbind(mixed_design,
                           mixed_design,
                           mixed_design,
                           mixed_design) %>%
  
  mutate(Subjects = rep(1:36, each = 3),
         recall = sample(0:20, 36*3, replace=TRUE))


full_mixed_design %>%
  group_by(OL, IL) %>%
  summarize(mean_recall = mean(recall), .groups="drop")

```



## References

















