---
title: "Lab3"
author: "Anna Matarazzo"
date: "3/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Problems

### Problem 1

The problem for this week is to develop a little bit of content (tutorial material) that could be added to this lab. (6 points)



### The difference between partial and semi-partial correlation

Partial correlation coefficient is a measure that is used to describe the relationship between two variables when BOTH of these variables are controlled (or partialled out) for the effects of a THIRD variable.


A semi-partial correlation describes the relationship between two variables when ONE of them is controlled for the effect of a THIRD variable.

Let's consider having three variables: X, Y, Z. The relationship of interest is between X and Y (so let's call them variables of interest). To illustrate the difference between types of correlation, let's illustrate the following scenarios:

1. None of the variables of interest are correlated with Z. 

PIC 1

2. Both of the variables of interest are correlated with Z.

PIC 2

3. Only one variable of interest (for example, Y) is correlated with Z.


```{r}

```







Since both X and Y has nothing to do with Z
This would be an example of linear correlation between two variables - X and Y.















