---
title: "Lab4"
author: "Anna Matarazzo"
date: "3/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(papaja)
library(ggplot2)
library(dplyr)
```



## Problems

### Problem 1

Consider the following example data for a between-subjects experiment with two groups, A and B:
```{r}
example_data <- tibble(Group = rep(c("A","B"), each = 5),
                       DV = c(2,4,3,5,4,7,6,5,6,7))
```
Use R to conduct a t.test and ANOVA on this data. Then use R to prove that the results of both analyses are the same. For example, prove that the p-values are the same, and prove that the F-value and T-value are related. (3 points)


### My version

```{r}

library(tibble)
example_data <- tibble(Group=rep(c("A","B"), each = 5), 
                                 DV=c(2,4,3,5,4,7,6,5,6,7))


table <- tibble(Test_type=rep(c("T-test", "Anova"), 2),
                              Results = c(
## p.value
t.test(DV~Group, data = example_data, var.equal=TRUE)$p.value,
summary(aov(DV~Group, data = example_data))[[1]]$`Pr(>F)`[1],

## F value
t.test(DV~Group, data = example_data, var.equal=TRUE)$statistic^2,
summary(aov(DV~Group, data = example_data))[[1]]$`F value`[1])
)

print(table)

```


### Key version

```{r}
my_t <- t.test(DV~Group, var.equal=TRUE,data=example_data)

my_aov <- summary(aov(DV~Group,data=example_data))

my_t$p.value

my_aov[[1]]$`Pr(>F)`[1]

round(my_t$p.value, digits =5) == round(my_aov[[1]]$`Pr(>F)`[1], digits=5)

my_t$statistic^2
my_aov[[1]]$`F value`[1]

round(my_t$statistic^2, digits = 1) == round(my_aov[[1]]$`F value`[1], digits = 1)

```


### Problem 2

  Look at the lab on ANOVA that I wrote for our undergraduate statistics OER lab manual https://crumplab.github.io/statisticsLab/lab-8-one-way-anova.html.    That lab shows an example of obtaining data from a published paper in psych science where a one-factor ANOVA was used as a part of the analysis. Load the data, conduct the ANOVA, report a ggplot of the means, and use papaja to help you write a short results section reporting the ANOVA result. (3 points).

```{r}
library(data.table)

all_data <- fread("data/Jamesetal2015Experiment2.csv")

all_data$Condition <- as.factor(all_data$Condition)
levels(all_data$Condition) <- c("Control", "React_Tetris", "Tetris", "React")

anova.out <- aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, data = all_data)
anova.out

summary((anova.out))


ggplot(all_data, aes(x=Condition, y=Days_One_to_Seven_Number_of_Intrusions))+
  geom_bar(stat="summary", fun="mean", position = "dodge")+
  geom_point()

```


We found a main effect of Intervention Type, `r apa_print(anova.out)$full_result$Condition`.













