---
title: "Lab9"
author: "Anna Matarazzo"
date: "4/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tibble)
library(tidyr)
library(dplyr)

e1 <- tribble(~Subject, ~Drug_A, ~Placebo, ~Drug_B,
        "s1", 124, 108, 104,
        "s2", 105, 107, 100,
        "s3",107,90,100,
        "s4",109,89,93,
        "s5",94,105,89,
        "s6",121,71,84)

# convert to long
e1 <- pivot_longer(e1,cols = !Subject,
                   names_to = "IV",
                   values_to = "Latency")

# Convert IVs to factors
e1 <- e1 %>%
  mutate(Subject = as.factor(Subject),
         IV = as.factor(IV))

# conduct ANOVA, note the addition of the error term in the formula
aov_out <- aov(Latency ~ IV + Error(Subject), data=e1)
summary(aov_out)

```


```{r}
e2 <- tribble(~Subject, ~r1, ~r2, ~r3, ~r4, ~r5, ~r6,
        "s1",30,18,21,15,18,12,
        "s2",21,23,16,17,13,12,
        "s3",19,21,13,15,13,9,
        "s4",19,19,16,9,11,10,
        "s5",21,16,12,15,9,11,
        "s6",22,17,14,12,10,9,
        "s7",19,20,17,10,13,5,
        "s8",17,18,11,11,9,12
        )

# convert to long
e2 <- pivot_longer(e2,cols = !Subject,
                   names_to = "Rank",
                   values_to = "Recall")

# Convert IVs to factors
e2 <- e2 %>%
  mutate(Subject = as.factor(Subject),
         Rank = as.factor(Rank))

# conduct ANOVA, note the addition of the error term in the formula
aov_out <- aov(Recall ~ Rank + Error(Subject), data= e2)
summary(aov_out)

```

```{r}
godden_baddeley <- tribble(~Subjects,~LearningPlace,~TestingPlace,~Recall,
        "s1","On Land","On Land",34,
        "s2","On Land","On Land",37,
        "s3","On Land","On Land",27,
        "s4","On Land","On Land",43,
        "s5","On Land","On Land",44,
        "s1","On Land","Under Sea",18,
        "s2","On Land","Under Sea",21,
        "s3","On Land","Under Sea",25,
        "s4","On Land","Under Sea",37,
        "s5","On Land","Under Sea",34,
        "s1","Under Sea","On Land",14,
        "s2","Under Sea","On Land",21,
        "s3","Under Sea","On Land",31,
        "s4","Under Sea","On Land",27,
        "s5","Under Sea","On Land",32,
        "s1","Under Sea","Under Sea",22,
        "s2","Under Sea","Under Sea",25,
        "s3","Under Sea","Under Sea",33,
        "s4","Under Sea","Under Sea",33,
        "s5","Under Sea","Under Sea",42
        )

# convert IVs to factors
godden_baddeley <- godden_baddeley %>%
  mutate(Subjects = as.factor(Subjects),
         LearningPlace = as.factor(LearningPlace),
         TestingPlace = as.factor(TestingPlace))

# run ANOVA
aov_out <- aov(Recall ~ LearningPlace*TestingPlace + Error(Subjects/(LearningPlace*TestingPlace)), godden_baddeley)

# print out ANOVA summary table
summary(aov_out)

# generate plot of means
library(ggplot2)

ggplot(godden_baddeley, aes(x=TestingPlace,
                            y=Recall,
                            shape=LearningPlace,
                            group=LearningPlace))+
  geom_point(stat="summary",fun="mean")+
  geom_line(stat="summary",fun="mean")+
  theme_classic(base_size=12)
```

```{r}
## example data set to illustrate sphericity
textbook <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,46,30,
        "s3",58,34,32,28,
        "s4",46,46,32,28,
        "s5",30,18,36,28
        )

## selecting the table of values in each column
textbook[,2:5]

## computing the covariation matrix
cov(textbook[,2:5])

## ta
colMeans(cov(textbook[,2:5]))

## sa-s
colMeans(cov(textbook[,2:5])) - mean(cov(textbook[,2:5]))

## double-centering

cov_matrix <- cov(textbook[,2:5])
col_mean_matrix <- cov_matrix*0 + colMeans(cov_matrix)
row_mean_matrix <- t(cov_matrix*0 + rowMeans(cov_matrix))
dc_matrix <- cov_matrix - col_mean_matrix -row_mean_matrix + mean(cov_matrix)

## greenhouse-geisser
sum(diag(dc_matrix))^2 / ((dim(dc_matrix)[1]-1)*sum(dc_matrix^2))
#> [1] 0.4459613
```

```{r}

textbook <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,46,30,
        "s3",58,34,32,28,
        "s4",46,46,32,28,
        "s5",30,18,36,28
        )

## ANOVA using car
wide_data <- as.matrix(textbook[1:5,2:5])
aov_model <- lm(wide_data ~1)
rm_factor <- factor(c('a1','a2','a3','a4'))
library(car)
new_anova <- Anova(aov_model,
                   idata=data.frame(rm_factor),
                   idesign = ~rm_factor,
                   type="III")
summary(new_anova,multivariate=FALSE)

```



Problems
Create an R script that can generate simulated data for the following repeated measures design. (2 points)
A. The dependent variable is assumed to come from a normal distribution with mean = 0 and standard deviation = 1.

B. There is one repeated measures factor with 5 levels (Down2, Down1, Control, Up1, Up2). The control group is assumed to have no effect. The Down1 and Down2 levels shift the mean down by 1 and 2 standard deviations, respectively. The Up1 and Up2 levels shift the mean up by 1 and 2 standard deviations, respectively.

C. There are 6 subjects in the experiment, and they are each measured once in each condition. The 6 subjects are assumed to be different from one another (e.g., they will have different baseline means in the control condition), but they will all be influenced by the IV in the exact same way (e.g., no interaction).


```{r}
# mean=0, sd=1

library(tibble)

sim_data <- tibble(
subjects = rep(1:6, each=5),
IV = rep(c("Down2", "Down1","Control","Up1","Up2"), each=6),
DV = 0
)

sim_data <- tibble(
subjects = rep(1:6, each=5),
IV = rep(c("Down2", "Down1","Control","Up1","Up2"), each=6),
DV = c(0,0,0,0,0,
       0.5, 0.5, 0.5, 0.5, 0.5,
        -0.25,-0.25,-0.25,-0.25,-0.25,
       -2,-2,-2,-2,-2,
       -1.5,-1.5,-1.5,-1.5,-1.5,
       1.5, 1.5, 1.5, 1.5, 1.5)
)

ggplot(sim_data, aes(x=IV, y=DV, group=subjects))+
  geom_point()+
  geom_line()
```
Run a simulation to determine the proportion of experiments that would return a significant result for the above design. Assume alpha = .05. (2 points)

Demonstrate that the Godden and Baddeley example data from the textbook (19.5), which used a 2x2 repeated measures design, can be be analyzed with one-sample t-tests to return the same results. Specifically, show the one-sample t-tests for each main effect and the interaction. (2 points)



```{r}

```
Bonus Points

These bonus questions are about the sphericity concept, and they will involve modifying the example data from the textbook (same as used in concept sectdion 1). The data is reprinted here for your convenience:

sphericity <- tribble(~S, ~a1, ~a2, ~a3, ~a4,
        "s1",76,64,34,26,
        "s2",60,48,46,30,
        "s3",58,34,32,28,
        "s4",46,46,32,28,
        "s5",30,18,36,28
        )
Create a line plot showing how each of the 5 subjects perform across the levels of the IV. Discuss how the line plot visually shows the sphericity problem in the data (1 point)

Modify the above data to remove the sphericity problem. Specifically, ensure that all subjects are different from another (there overall means are different), and that the IV has the same effect for each level and each subject (no interaction). Then, plot the new data and discuss how the graph shows the sphericity problem has been removed. (1 point)

Calculate the Greenhouse-Geisser estimate of epsilon for your modified data to demonstrate you have removed the sphericity problem. (1 point)


